New insights into the Weddell Sea ecosystem applying a network approach

### 2022-08-10 Manuscript/Manuscript.Rmd
M&M, Results: 1st draft finished

We analysed the food web of the Weddell Sea by: a) estimating the strength of each interaction; b) studying the properties of the species in a network approach; and c) comparing the stability of the food web after performing extinction simulations.

@@

To estimate the strength of each interaction in the food web, we followed the methodology proposed by @Pawar2012. The minimum data requirements are: body mass of the consumer (predator) and resource (prey), and the interaction dimensionality classified as 2 or 3 dimensions. GATEWAy v.1.0 does provide information on the mean mass for consumers and resources (except for 'detritus' and 'sediment') for every interaction, but lacks the dimensionality for 924 interactions. To solve this issue, we used the information about movement type for consumer and resource. Then, we classified the interaction as 2D when both consumer and resource move in 2D (e.g., both are sessile or walking) or if a consumer moves in 3D and a resource in 2D (e.g., swimming consumer and sessile/walking resource). The interaction was classified as 3D when both consumer and resource move in 3D (e.g., both swimming) or if the consumer moves in 2D and resource in 3D (e.g., sessile/walking consumer, swimming resource) [@Pawar2012].

The main equation we used for estimating the interaction strength IS is:

@@

where \vec{\alpha} is the search rate, \vec{x_R} is the density of the resource, and \vec{m_R} and \vec{m_C} are the body mass of the resource and the consumer, respectively [@Pawar2012].

We obtained estimations for the resource density and the search rate from the scaling relationships with the resource and the consumer mass, respectively (@Pawar2012). The coefficients of such relationships, determined by ordinary least squares regression, vary with the interaction dimensionality. On one hand, resource density scales with resource mass as a power-law with exponents \vec{p = -0.79 \pm 0.09} in 2D and \vec{p = -0.86 \pm 0.06} in 3D. Since mean mass for resources 'detritus' and 'sediment' were not available in GATEWAy v.1.0, we calculated it considering the scaling relationship with consumer mass, also different in 2D and 3D (for details see equation S9 and figures 2c-d of Supplementary Information in @Pawar2012). On the other hand, search rate scales with consumer mass as a power-law with exponents \vec{p = 0.68 \pm 0.12} in 2D and \vec{p = 1.05 \pm 0.08} in 3D.

Finally, we fit the distribution of the interaction strengths of the food web considering six candidate models (Uniform, Normal, Exponential, Power-law, log-Normal and Gamma) using maximum likelihood [@McCallum2008], and selected the model performance by computing the Akaike Information Criterion [@Burnham2002].

@@

As weighted property we took into account the mean interaction strength, meaning the average strength of all interactions. After exploring the distribution of the species log(mean interaction strength), we decided to apply the k-means clustering method, which aims to partition the species into k groups such that the sum of squares from points to the assigned cluster centres is minimized [@Hartigan1979]. Then we tested the goodness of this clustering measure with the Gap statistic, which compares log(W(k)) with E*[log(W(k))] where the latter is defined via bootstraping [@Tibshirani2001].

On the other hand, we considered the following unweighted properties: a) degree or the total number of trophic interactions, taking into account in- and out-interactions (role as predator and prey, respectively); b) trophic level or the position in the food web relative to primary producers/detritus; and c) trophic similarity or the trophic overlap between species based on shared and unique resources and consumers.

>>>>

We analysed the food web of the Weddell Sea by: a) estimating the strength of each interaction; b) studying the properties of the species in a network approach; and c) comparing the stability of the food web after performing species extinction simulations.

@@

To estimate the strength of each interaction in the food web, we followed the methodology proposed by @Pawar2012. The minimum data requirements are: body mass of the consumer (predator) and resource (prey), and the interaction dimensionality classified as 2 or 3 dimensions. GATEWAy v.1.0 does provide the mean mass for consumers and resources (except for 'detritus' and 'sediment') and the dimensionality for every interaction, though the latter is missing for 924 interactions. To solve this issue, we used information about movement type for consumer and resource included in GATEWAy. Then, we classified the interaction as 2D when both consumer and resource move in 2D (e.g., both are sessile or walking) or if a consumer moves in 3D and a resource in 2D (e.g., swimming consumer and sessile/walking resource). The interaction was classified as 3D when both consumer and resource move in 3D (e.g., both swimming) or if the consumer moves in 2D and resource in 3D (e.g., sessile/walking consumer, swimming resource) [@Pawar2012].

The main equation we used for estimating the interaction strength IS was:

@@

where \vec{\alpha} is the search rate, \vec{x_R} is the resource density, and \vec{m_R} and \vec{m_C} are the body mass of the resource and the consumer, respectively [@Pawar2012].

We obtained estimations for the resource density and the search rate from the scaling relationships with the resource and the consumer mass, respectively [@Pawar2012]. The coefficients of such relationships, determined by ordinary least squares regression, vary with the interaction dimensionality. On one hand, resource density scales with resource mass as a power-law with exponents \vec{p = -0.79 \pm 0.09} in 2D and \vec{p = -0.86 \pm 0.06} in 3D. Since mean mass for resources 'detritus' and 'sediment' were not available in GATEWAy, we estimated it considering the scaling relationship with consumer mass, also different in 2D and 3D (for details see equation S9 and figures 2c-d of Supplementary Information in @Pawar2012). On the other hand, search rate scales with consumer mass as a power-law with exponents \vec{p = 0.68 \pm 0.12} in 2D and \vec{p = 1.05 \pm 0.08} in 3D.

After this, we fit the distribution of the interaction strengths of the food web considering six candidate models (Uniform, Normal, Exponential, Power-law, log-Normal and Gamma) using maximum likelihood [@McCallum2008], and selected the model performance by computing the Akaike Information Criterion AIC [@Burnham2002].

@@

As the weighted property we took into account the mean interaction strength, meaning the average strength of all interactions. After exploring the distribution of the species' log(mean interaction strength), we decided to apply the k-means clustering method, which aims to partition the species into k groups such that the sum of squares from points to the assigned cluster centres is minimized [@Hartigan1979]. Then we tested the goodness of this clustering measure with the Gap statistic, which is a kind of data mining algorithm to improve the clustering process by efficient estimation of the best number of clusters [@Tibshirani2001].

On the other hand, we considered the following unweighted properties: a) degree or the total number of trophic interactions, taking into account in- and out-coming interactions (role as predator and prey, respectively); b) trophic level or the position in the food web relative to primary producers/detritus; and c) trophic similarity or the trophic overlap between species based on shared and unique resources and consumers. The following are arguments to have selected the above properties. The degree has often been equated with importance to the structure and functioning of a community, i.e. perturbations to high-degree species may therefore have larger effects on the food web than perturbations to low-degree species [@Dunne2002a; references in @Cirtwill2018a]. The trophic level offers information about how important a species is to its biotic community, i.e. top predators and primary producers are expected to have particularly large effects on the rest of their communities through top-down and bottom-up control, respectively [references in @Cirtwill2018a]. The trophic similarity measures the similarity of one of the most important aspect of species’ niches, the trophic niche, and functional aspects of biodiversity [@Martinez1991; @Williams2000].

### 2022-08-10 Manuscript/Manuscript.tex
M&M, Results: 1st draft finished

To estimate the strength of each interaction in the food web, we
followed the methodology proposed by \citet{Pawar2012}. The minimum data
requirements are: body mass of the consumer (predator) and resource
(prey), and the interaction dimensionality classified as 2 or 3
dimensions. GATEWAy v.1.0 does provide information on the mean mass for
consumers and resources (except for `detritus' and `sediment') for every
interaction, but lacks the dimensionality for 924 interactions. To solve
this issue, we used the information about movement type for consumer and
resource. Then, we classified the interaction as 2D when both consumer
and resource move in 2D (e.g., both are sessile or walking) or if a
consumer moves in 3D and a resource in 2D (e.g., swimming consumer and
sessile/walking resource). The interaction was classified as 3D when
both consumer and resource move in 3D (e.g., both swimming) or if the
consumer moves in 2D and resource in 3D (e.g., sessile/walking consumer,
swimming resource) \citep{Pawar2012}.

The main equation we used for estimating the interaction strength IS is:

@@

where \vec{\alpha} is the search rate, \vec{x_R} is the density of the
resource, and \vec{m_R} and \vec{m_C} are the body mass of the resource
and the consumer, respectively \citep{Pawar2012}.

We obtained estimations for the resource density and the search rate
from the scaling relationships with the resource and the consumer mass,
respectively (\citet{Pawar2012}). The coefficients of such
relationships, determined by ordinary least squares regression, vary
with the interaction dimensionality. On one hand, resource density
scales with resource mass as a power-law with exponents
\vec{p = -0.79 \pm 0.09} in 2D and \vec{p = -0.86 \pm 0.06} in 3D. Since
mean mass for resources `detritus' and `sediment' were not available in
GATEWAy v.1.0, we calculated it considering the scaling relationship
with consumer mass, also different in 2D and 3D (for details see
equation S9 and figures 2c-d of Supplementary Information in
\citet{Pawar2012}). On the other hand, search rate scales with consumer
mass as a power-law with exponents \vec{p = 0.68 \pm 0.12} in 2D and
\vec{p = 1.05 \pm 0.08} in 3D.

Finally, we fit the distribution of the interaction strengths of the
food web considering six candidate models (Uniform, Normal, Exponential,
Power-law, log-Normal and Gamma) using maximum likelihood
\citep{McCallum2008}, and selected the model performance by computing
the Akaike Information Criterion \citep{Burnham2002}.

@@

As weighted property we took into account the mean interaction strength,
meaning the average strength of all interactions. After exploring the
distribution of the species log(mean interaction strength), we decided
to apply the k-means clustering method, which aims to partition the
species into k groups such that the sum of squares from points to the
assigned cluster centres is minimized \citep{Hartigan1979}. Then we
tested the goodness of this clustering measure with the Gap statistic,
which compares log(W(k)) with E*{[}log(W(k)){]} where the latter is
defined via bootstraping \citep{Tibshirani2001}.

On the other hand, we considered the following unweighted properties: a)
degree or the total number of trophic interactions, taking into account
in- and out-interactions (role as predator and prey, respectively); b)
trophic level or the position in the food web relative to primary
producers/detritus; and c) trophic similarity or the trophic overlap
between species based on shared and unique resources and consumers.

>>>>

To estimate the strength of each interaction in the food web, we
followed the methodology proposed by \citet{Pawar2012}. The minimum data
requirements are: body mass of the consumer (predator) and resource
(prey), and the interaction dimensionality classified as 2 or 3
dimensions. GATEWAy v.1.0 does provide the mean mass for consumers and
resources (except for `detritus' and `sediment') and the dimensionality
for every interaction, though the latter is missing for 924
interactions. To solve this issue, we used information about movement
type for consumer and resource included in GATEWAy. Then, we classified
the interaction as 2D when both consumer and resource move in 2D (e.g.,
both are sessile or walking) or if a consumer moves in 3D and a resource
in 2D (e.g., swimming consumer and sessile/walking resource). The
interaction was classified as 3D when both consumer and resource move in
3D (e.g., both swimming) or if the consumer moves in 2D and resource in
3D (e.g., sessile/walking consumer, swimming resource)
\citep{Pawar2012}.

The main equation we used for estimating the interaction strength IS
was:

@@

where \vec{\alpha} is the search rate, \vec{x_R} is the resource
density, and \vec{m_R} and \vec{m_C} are the body mass of the resource
and the consumer, respectively \citep{Pawar2012}.

We obtained estimations for the resource density and the search rate
from the scaling relationships with the resource and the consumer mass,
respectively \citep{Pawar2012}. The coefficients of such relationships,
determined by ordinary least squares regression, vary with the
interaction dimensionality. On one hand, resource density scales with
resource mass as a power-law with exponents \vec{p = -0.79 \pm 0.09} in
2D and \vec{p = -0.86 \pm 0.06} in 3D. Since mean mass for resources
`detritus' and `sediment' were not available in GATEWAy, we estimated it
considering the scaling relationship with consumer mass, also different
in 2D and 3D (for details see equation S9 and figures 2c-d of
Supplementary Information in \citet{Pawar2012}). On the other hand,
search rate scales with consumer mass as a power-law with exponents
\vec{p = 0.68 \pm 0.12} in 2D and \vec{p = 1.05 \pm 0.08} in 3D.

After this, we fit the distribution of the interaction strengths of the
food web considering six candidate models (Uniform, Normal, Exponential,
Power-law, log-Normal and Gamma) using maximum likelihood
\citep{McCallum2008}, and selected the model performance by computing
the Akaike Information Criterion AIC \citep{Burnham2002}.

@@

As the weighted property we took into account the mean interaction
strength, meaning the average strength of all interactions. After
exploring the distribution of the species' log(mean interaction
strength), we decided to apply the k-means clustering method, which aims
to partition the species into k groups such that the sum of squares from
points to the assigned cluster centres is minimized
\citep{Hartigan1979}. Then we tested the goodness of this clustering
measure with the Gap statistic, which is a kind of data mining algorithm
to improve the clustering process by efficient estimation of the best
number of clusters \citep{Tibshirani2001}.

On the other hand, we considered the following unweighted properties: a)
degree or the total number of trophic interactions, taking into account
in- and out-coming interactions (role as predator and prey,
respectively); b) trophic level or the position in the food web relative
to primary producers/detritus; and c) trophic similarity or the trophic
overlap between species based on shared and unique resources and
consumers. The following are arguments to have selected the above
properties. The degree has often been equated with importance to the
structure and functioning of a community, i.e.~perturbations to
high-degree species may therefore have larger effects on the food web
than perturbations to low-degree species
\citetext{\citealp{Dunne2002a}; \citealp[references
in][]{Cirtwill2018a}}. The trophic level offers information about how
important a species is to its biotic community, i.e.~top predators and
primary producers are expected to have particularly large effects on the
rest of their communities through top-down and bottom-up control,
respectively \citep[references in][]{Cirtwill2018a}. The trophic
similarity measures the similarity of one of the most important aspect
of species' niches, the trophic niche, and functional aspects of
biodiversity \citep{Martinez1991, Williams2000}.
